 

### 2023-05-11 paper2.Rmd
Final revisions to manuscript

{\bf Findings and interpretation:} We found a statistically significant reduction in monthly visits reported associated with temperature—approximately `r function_dynamic_text_results(section = "main", coefficient_option = "exposure_temperature", model_option = 4, value = "irr", value_include = "only")` visits reported per each degree Celsius increase `r function_dynamic_text_results(section = "main", coefficient_option = "exposure_temperature", model_option = 4, value = "irr", value_include = TRUE)`. A multiverse analysis of `r num_models_multiverse` models found this estimate to be robust to alternative specifications; ecoregion analysis indicated that this association was primarily localized to facilities in the central highlands, where cooler temperatures are more prevalent. We further found that higher soil moisture may be associated with higher mental health related visits in the future, with an `r function_dynamic_text_results(section = "lags", lag_option = 3, coefficient_option = "exposure_sm", value = "irr", value_include = "only")` elevation in visits reported `r function_dynamic_text_results(section = "lags", lag_option = 3, coefficient_option = "exposure_sm", value = "irr", value_include =TRUE)` after a three month lag. Ecoregion analysis suggests more visits are reported in the eastern lowlands and the northern dry deciduous forests when soil moisture rises, which is consistent with the geographic concentration of flood risk in Madagascar. Our findings highlight the complex interactions between CCEs and population mental health and suggest potential heterogeneity in how changing climate conditions will impact mental health needs and system responses.

@@

Among low-income countries most vulnerable to the health consequences of climate change is Madagascar, a nation that generates one percent *of* one percent of global carbon dioxide emissions. Worsening \acr{CCEs} in Madagascar (primarily in the form of droughts, floods, and cyclones) are likely to exacerbate endemic, climate-sensitive threats to health, including malnutrition and malaria.[@rakotoarisonAssessmentRiskVulnerability2018] In particular, persistent drought conditions in recent years have led to extensive food insecurity.[@adamtaylorMadagascarHeadedClimate] Though there is debate concerning the extent to which human activity is directly responsible for Madagascar's current drought,[@harringtonAttributionSevereLow; @zhongManyDimensionsDrought2021] the extremity of the situation has led some observers to classify conditions in Madagascar as the first climate change famine in history.[@worldfoodprogrammeMadagascarSevereDrought2021]

@@

Among facilities reporting any mental health data between 2010 and 2020, data were obtained from `r num_facilities_CSB` public and private primary health centers, known as Centres Santé de Base or CSBs and Formations Sanitaires Privées de Base or FSBs, respectively), `r num_facilities_CHD` district referral hospitals (CHDs), `r num_facilities_CHRR_CHU` regional referral hospitals (CHRRs) and university hospitals (CHUs), for a total of `r num_facilities_pretty` uniquely-identified facilities in our sample.

Exploratory data analysis indicated that the classification and reporting of counts of mental disorders in \acr{GESIS} varied across types of facilities in their level of detail. Among CHD facilities, incidence of mental disorders was aggregated with neurological disorders (coded as “neuro psychiatric disorders”), while CHRR and CHU facilities reported more detailed counts of specific mental disorders, such as depression, personality disorders, and schizophrenia. Across all facilities, disaggregated data was reported by age group but not by sex. Among CSB facilities, incidences of mental disorders were reported by age group under “mental illnesses and psychic disorders” until June 2015, when reporting was reclassified to “mental disorders,” and disaggregated data was reported by both age and sex. A summary of classification by facility type and year is provided in Appendix 2.1, Table S2.1, with English translations of the French disease names provided in Table S2.2. and yearly cumulative reported visit totals by disorder reported by facility level and year in Tables S2.3-5.

To address heterogeneity in reporting, our analysis aggregated any monthly visits concerning mental or neurological disorders reported by a facility, irrespective of age and sex, into a summary outcome measure. To address outliers, monthly observations of the outcome were arranged separately for each facility into a unique time series. The nonparametric Friedman’s super smoother regression estimator[@friedmanVariableSpanScatterplot1984] was applied to each time series to identify outlier observations and replace them through linear interpolation.[@hyndmanAutomaticTimeSeries2008] By considering outliers on a facility-by-facility basis, the data cleaning approach preserved extreme values that may reflect true variances in incidence (for example, incidences reported by a large, regional reference hospital) while addressing potential errors in data entry. (Additional detail on the data cleaning process is available in Appendix 2.1)

@@

To retrieve spatial climate data for the facilities in our sample, it was first necessary to geocode facilities to obtain their latitude and longitude. Facility data from \acr{GESIS} included limited geospatial information. While exact facility coordinates were not available, information on the relevant administrative unit served by a facility was included in the data (e.g., regions for CHRRs, districts for CHDs, and communes for CSBs). These data informed a stepwise matching process to locate facilities in our sample.

@@

First, we considered the role of data cleaning in the outcome data and chose to compare our analysis using the cleaned outcome data against the raw, unadjusted values. Second, we chose to compare our findings using data from all facilities (the full sample) against a sample without district and regional hospital (CSB/FSP facilities only). Primary care facilities have smaller catchment areas than hospitals; consequently, it is more plausible that climate data for primary care facilities will closely reflect climate exposure to patients compared to referral hospitals. Third, we considered the decision to model temperature and soil moisture as an additive or interactive relationship (Model 3 vs. Model 4), with both alternatives included in our multiverse analysis. Fourth, we evaluated our decision to account for overdispersion with a \acr{NB} model by using a Poisson regression framework as an alternative. Fifth, we evaluated alternative specifications of exposure to cyclone activity, using three options: our binary indicator, our count indicator, and no term for cyclone activity included. Sixth, we varied whether time fixed effects (month and year) were included. Seventh, we similarly included and excluded regional fixed effects for comparison.

@@

We then interacted the methodological choices to generate `r num_models_multiverse` plausible models. Using the effect estimates and confidence intervals for these models, we constructed specification curves for monthly temperature, soil moisture, and—for the `r as.numeric(num_models_multiverse)/3` models that included the cyclone count indicator— cyclone activity to visually compare the distribution of plausible estimates under different combinations.

@@

Figure \ref{figure_gps_climate} shows the spatial distribution of temperature and soil moisture across the facilities in our sample during January and June 2015, corresponding to the warmer, wet season and the dryer, cool season, respectively. When evaluated over time, annual climate trends in Madagascar showed relatively stable distribution of average temperatures, while soil moisture had a slight upward shift through 2020 (Figure \ref{figure_climate_distribution}). Data of tropical storm and cyclone activity throughout the study period also exhibited seasonal patterns, with most major storms occurring in the hot season, with a peak between February and May (Figure \ref{figure_cyclone}).

@@

```{r figure_gps_climate, include = TRUE, echo = FALSE, warnings = FALSE, fig.scap="Spatial distribution of climate conditions across clinics and seasons in Madagascar.", out.extra='', fig.cap="\\label{figure_gps_climate}Spatial distribution of climate conditions across clinics and seasons in Madagascar. Mean monthly temperature (top) is measured in degrees Celcius and mean soil moisture is measured in percent saturation (below).", fig.width = 6, fig.height = 7.5, fig.align='center', out.width = '90%'}

@@

The regression models predicting monthly counts of any mental or neurological disorder (Table 2.2) showed significant associations with temperature across all models. A one-degree Celsius increase in temperature was associated with approximately `r function_dynamic_text_results(section = "main", model_option = 1, coefficient_option = "exposure_temperature", value = "irr", value_include ="only")` visits reported `r function_dynamic_text_results(section = "main", model_option = 1, coefficient_option = "exposure_temperature", value = "irr", value_include =TRUE)` by primary clinics and hospitals in Model 1. Inclusion of both climate predictors resulted in little change to the measure of association for temperature. In an additive model with soil moisture included (Model 3), the effect estimate associated with temperature grew to `r function_dynamic_text_results(section = "main", model_option = 3, coefficient_option = "exposure_temperature", value = "irr", value_include ="only")` visits reported `r function_dynamic_text_results(section = "main", model_option = 3, coefficient_option = "exposure_temperature", value = "irr", value_include =TRUE)`. Adopting an interactive model (Model 4) attenuated the association `r function_dynamic_text_results(section = "main", model_option = 4, coefficient_option = "exposure_temperature", value = "irr", value_include =TRUE)`, but remained significant at a 95% level of confidence. Across all models with temperature, the average marginal effect associated with temperature (across a sample with all clinics) ranged between 2 and 3 fewer monthly mental health visits reported and was statistically significant at an alpha of 0.01, with the average marginal effect in Model 4 being `r function_dynamic_text_results(section = "main", model_option = 4, coefficient_option = "exposure_temperature", value = "ame", value_include ="only")` visits `r function_dynamic_text_results(section = "main", model_option = 4, coefficient_option = "exposure_temperature", value = "ame", value_include =FALSE)`.

@@

```
Predicted values of monthly mental health related visits counts under Model 4 echoed the effect estimates from Table 2, with fewer visits reported in warmer months (Figure \ref{figure_marginal}). Marginal effect estimates across the distribution of temperatures in our data showed that the greatest level of reduction occurred between 10 and 15 degrees Celsius, with each degree increase associated with 4 to 5 fewer visits related to mental or neurological disorders reported, controlling for other covariates. Months with greater levels of soil moisture and days with cyclone activity also showed fewer reported visits; however, the marginal effects associated with soil moisture were not significant across any unit of exposure, while those associated with days with cyclone activity were close to zero for all lengths of activity.

Examining lagged effects showed potential temporal associations between soil moisture levels in preceding months and visit counts in those following (Figure \ref{figure_lag}), with higher saturation leading to a statistically significant elevation in visits reported 1 to 5 months later. At a lag of 3 months, the elevation in visits was approximately `r function_dynamic_text_results(section = "lags", lag_option = 3, coefficient_option = "exposure_sm", value = "irr", value_include = "only")` `r function_dynamic_text_results(section = "lags", lag_option = 3, coefficient_option = "exposure_sm", value = "irr", value_include =TRUE)`. There appeared to be a significant elevation of similar magnitude 13 to 17 months prior, suggesting potential annularity in lagged association. Effects for temperature were not significant across any lag; cyclone activity similarly showed no significance in lags under a year, but did show modest, statistically significant associations at lags of 12 months `r function_dynamic_text_results(section = "lags", lag_option = 12, coefficient_option = "exposure_cyclone", value = "irr", value_include = TRUE)` and 15 months `r function_dynamic_text_results(section = "lags", lag_option = 15, coefficient_option = "exposure_cyclone", value = "irr", value_include = TRUE)`. A slight increase in visits was associated with cyclone activity in a prior month, but this increase was not statistically significant `r function_dynamic_text_results(section = "lags", lag_option = 1, coefficient_option = "exposure_cyclone", value = "irr", value_include = TRUE)`.

@@

```{r figure_sensitivity_analysis_temp, include = TRUE, echo=FALSE, warnings = FALSE, fig.scap=paste("Specification curve depicting the effect estimates and confidence intervals for mean monthly temperature (degrees Celsius) across", num_models_multiverse, "models."), out.extra='', fig.cap=paste("\\label{figure_sensitivity_analysis_temp}Specification curve depicting the effect estimates and confidence intervals for mean monthly temperature (degrees Celsius) across", num_models_multiverse, "models. In the top panel, estimated incidence rate ratios (IRRs) across different models are depicted as circles; the darkly and lightly shaded regions above and below these circles correspond to 95\\% and 99\\% confidence intervals of these estimates, respectively. The dotted horizontal line depicts a null effect. The IRR estimate and confidence intervals of the preferred model, Model 4, are highlighted in orange. Below each model, a legend of shaded and unshaded boxes indicates which modeling decisions correspond to a given model, with boxes shaded teal showing active features. FE: fixed effects. CSB/FSP: Centres Santé de Base and Formations Sanitaires Privées de Base. NB: negative binomial model.", sep=" "), fig.width = 15, fig.height = 9, fig.align='center', out.width = '95%'}

@@

```{r figure_sensitivity_analysis_sm, include = TRUE, echo=FALSE, warnings = FALSE, fig.scap=paste("Specification curve depicting the effect estimates and confidence intervals for mean monthly soil moisture (percent saturation) across", num_models_multiverse, "models."), out.extra='', fig.cap=paste("\\label{figure_sensitivity_analysis_sm}Specification curve depicting the effect estimates and confidence intervals for mean monthly soil moisture (percent saturation) across", num_models_multiverse, "models. In the top panel, estimated incidence rate ratios (IRRs) across different models are depicted as circles; the darkly and lightly shaded regions above and below these circles correspond to 95\\% and 99\\% confidence intervals of these estimates, respectively. The dotted horizontal line depicts a null effect. The IRR estimate and confidence intervals of the preferred model, Model 4, are highlighted in orange. Below each model, a legend of shaded and unshaded boxes indicates which modeling decisions correspond to a given model, with boxes shaded teal showing active features. FE: fixed effects. CSB/FSP: Centres Santé de Base and Formations Sanitaires Privées de Base. NB: negative binomial model.", sep=" "), fig.width = 15, fig.height = 9, fig.align='center', out.width = '95%'}

@@

```{r figure_sensitivity_analysis_cyclone, include = TRUE, echo=FALSE, warnings = FALSE, fig.width = 15, fig.height = 9, fig.align='center', out.width = '95%', fig.scap=paste("Specification curve depicting the effect estimates and confidence intervals for duration of cyclone activity across", num_models_multiverse, "models."), out.extra='', fig.cap=paste("\\label{figure_sensitivity_analysis_cyclone}Specification curve depicting the effect estimates and confidence intervals for cyclone activity (measured as days with a tropical storm or cyclone impacting Madagascar in a month) across", as.numeric(num_models_multiverse)/3, "models. In the top panel, estimated incidence rate ratios (IRRs) across different models are depicted as circles; the darkly and lightly shaded regions above and below these circles correspond to 95\\% and 99\\% confidence intervals of these estimates, respectively. The dotted horizontal line depicts a null effect. The IRR estimate and confidence intervals of the preferred model, Model 4, are highlighted in orange. Below each model, a legend of shaded and unshaded boxes indicates which modeling decisions correspond to a given model, with boxes shaded teal showing active features. FE: fixed effects. CSB/FSP: Centres Santé de Base and Formations Sanitaires Privées de Base. NB: negative binomial model.", sep=" ")}

@@

It is important to note in our interpretation of these findings that the true relationship between reported mental health visits and actual cases of mental disorders is not well understood, particularly in Madagascar, where there is no epidemiological data on the prevalence or distribution of mental disorders in the general population. The changes in mental health visits could reflect changes in the actual cases of mental disorders but could also result from the influence of \acr{CCEs} on help-seeking behavior, diagnostic capacity, or other factors. Additionally, it has been widely recognized that the majority individuals with mental disorders often do not receive any mental health treatment, particularly in low-income settings. [@demyttenaerePrevalenceSeverityUnmet2004]

@@

An alternative explanation could be that access to mental health services is limited in warmer months due to factors such as increased demand for medical services related to heat-related illnesses. A crowding out of provider availability due to physical complaints in warmer months (for example, malaria-related fever) could lead to decreased availability of provider access for mental disorders, especially at primary health facilities with already limited diagnostic capacity. Understanding how climate conditions in Madagascar impact the broader burden of disease, including physical conditions, is the subject of current investigation, and will lend further insight into potential mechanistic explanations for our findings.

@@

Though incidence rate ratios and marginal effect estimates associated with soil moisture were not statistically significant in our main analysis, we observed a significant association of higher soil moisture and visits for mental health complaints in the following quarter, echoing similar findings from United States meteorological data linking higher precipitation with worsened mental health.[@obradovichEmpiricalEvidenceMental2018] In addition, our subgroup analysis also indicated significant and diverging effects among different ecoregions. In particular, we found that greater soil moisture was associated with more visits reported in the eastern lowlands and the northern dry deciduous forests, which echoes the geographic concentration of flood risk in Madagascar.[@rakotoarisonSpatialAnalysisRisks2021] Conversely, the observed (though statistically modest) reduction in visits associated with higher soil moisture in the desert and xeric shrublands overlaps with the concentration of drought risk, where increased soil moisture could be taken as an indicator of reduced drought vulnerability. It is important to note, however, that the time frame of available GESIS data excludes the most severe period of the ongoing drought in Madagascar, limiting our ability to directly study the impact of the current crisis on population mental health in this analysis. Further research could leverage spatial reanalysis data on precipitation to further characterize these associations.

>>>>

{\bf Findings and interpretation:} We found a statistically significant reduction in monthly visits reported associated with temperature—approximately `r function_dynamic_text_results(section = "main", coefficient_option = "exposure_temperature", model_option = 4, value = "irr", value_include = "only")` visits reported per each degree Celsius increase `r function_dynamic_text_results(section = "main", coefficient_option = "exposure_temperature", model_option = 4, value = "irr", value_include = TRUE)`. A multiverse analysis of `r num_models_multiverse` models found this estimate to be robust to alternative specifications; ecoregion analysis indicated that this association was primarily localized to facilities in the central highlands, where cooler temperatures are more prevalent. We further found that higher soil moisture may be associated with higher mental health related visits in the future, with an `r function_dynamic_text_results(section = "lags", lag_option = 3, coefficient_option = "exposure_sm", value = "irr", value_include = "only")` elevation in visits reported `r function_dynamic_text_results(section = "lags", lag_option = 3, coefficient_option = "exposure_sm", value = "irr", value_include =TRUE)` after a three-month lag. Ecoregion analysis suggests more visits are reported in the eastern lowlands and the northern dry deciduous forests when soil moisture rises, which is consistent with the geographic concentration of flood risk in Madagascar. Our findings highlight the complex interactions between CCEs and population mental health and suggest potential heterogeneity in how changing climate conditions will impact mental health needs and system responses.

@@

Among low-income countries most vulnerable to the health consequences of climate change is Madagascar, a nation that generates one percent *of* one percent of global carbon dioxide emissions. Worsening \acr{CCEs} in Madagascar (primarily in the form of droughts, floods, and cyclones) are likely to exacerbate endemic, climate-sensitive threats to health, including malnutrition and malaria.[@rakotoarisonAssessmentRiskVulnerability2018] In particular, persistent drought conditions in recent years have led to extensive food insecurity.[@adamtaylorMadagascarHeadedClimate2021] Though there is debate concerning the extent to which human activity is directly responsible for Madagascar's current drought,[@harringtonAttributionSevereLow; @zhongManyDimensionsDrought2021] the extremity of the situation has led some observers to classify conditions in Madagascar as the first climate change famine in history.[@worldfoodprogrammeMadagascarSevereDrought2021]

@@

Among facilities reporting any mental health data between 2010 and 2020, data were obtained from `r num_facilities_CSB` public and private primary health centers, known as Centres Santé de Base or CSBs, and Formations Sanitaires Privées de Base or FSBs, respectively, `r num_facilities_CHD` district referral hospitals (CHDs), `r num_facilities_CHRR_CHU` regional referral hospitals (CHRRs) and university hospitals (CHUs), for a total of `r num_facilities_pretty` uniquely-identified facilities in our sample.

Exploratory data analysis indicated that the classification and reporting of counts of mental disorders in \acr{GESIS} varied across types of facilities in their level of detail. Among CHD facilities, incidence of mental disorders was aggregated with neurological disorders (coded as “neuro psychiatric disorders”), while CHRR and CHU facilities reported more detailed counts of specific mental disorders, such as depression, personality disorders, and schizophrenia. Across all facilities, disaggregated data were reported by age group but not by sex. Among CSB facilities, incidences of mental disorders were reported by age group under “mental illnesses and psychic disorders” until June 2015, when reporting was reclassified to “mental disorders”; disaggregated data were reported by both age and sex. A summary of classification by facility type and year is provided in Appendix 2.1, Table S2.1, with English translations of the French disease names provided in Table S2.2 and yearly cumulative reported visit totals by disorder reported by facility level and year in Tables S2.3-5.

To address heterogeneity in reporting, our analysis aggregated any monthly visits concerning mental or neurological disorders reported by a facility into a summary outcome measure, irrespective of age and sex. To address outliers, monthly observations of the outcome were arranged separately for each facility into a unique time series. The nonparametric Friedman’s super smoother regression estimator[@friedmanVariableSpanScatterplot1984] was applied to each time series to identify outlier observations and replace them through linear interpolation.[@hyndmanAutomaticTimeSeries2008] By considering outliers on a facility-by-facility basis, the data cleaning approach preserved extreme values that may reflect true variances in incidence (for example, incidences reported by a large, regional reference hospital) while addressing potential errors in data entry. (Additional detail on the data cleaning process is available in Appendix 2.1)

@@

To retrieve spatial climate data for the facilities in our sample, it was first necessary to geocode facilities to obtain their latitude and longitude. Facility data from \acr{GESIS} included limited information on the location of facilities. While exact facility coordinates were not available, information on the relevant administrative unit served by a facility was included in the data (e.g., regions for CHRRs, districts for CHDs, and communes for CSBs). These data informed a stepwise matching process to locate facilities in our sample.

@@

First, we considered the role of data cleaning of the outcome data and chose to compare our analysis using the cleaned outcome data against the raw unadjusted values. Second, we chose to compare our findings using data from all facilities (the full sample) against a sample without district and regional hospitals (CSB/FSP facilities only). Primary care facilities have smaller catchment areas than hospitals; consequently, it is more likely that climate data for primary care facilities will closely reflect patients' climate exposure compared to referral hospitals. Third, we considered the decision to model temperature and soil moisture as an additive or interactive relationship (Model 3 vs. Model 4), with both alternatives included in our multiverse analysis. Fourth, we evaluated our decision to account for overdispersion with a \acr{NB} model by using a Poisson regression framework as an alternative. Fifth, we evaluated alternative specifications of exposure to cyclone activity, using three options: our binary indicator, our count indicator, and no term for cyclone activity included. Sixth, we varied whether time fixed effects (month and year) were included. Seventh, we similarly included and excluded regional fixed effects for comparison.

@@

We then interacted the methodological choices to generate `r num_models_multiverse` plausible model specifications. Using the effect estimates and confidence intervals for these models, we constructed specification curves for monthly temperature, soil moisture, and—for the `r as.numeric(num_models_multiverse)/3` models that included the cyclone count indicator— cyclone activity to visually compare the distribution of plausible estimates for different combinations.

@@

Figure \ref{figure_gps_climate} shows the spatial distribution of temperature and soil moisture across the facilities in our sample during January and June 2015, corresponding to the warmer, wet season and the drier, cool season, respectively. When evaluated over time, annual climate trends in Madagascar showed a relatively stable distribution of average temperatures, while soil moisture had a slight upward shift through 2020 (Figure \ref{figure_climate_distribution}). Data of tropical storm and cyclone activity throughout the study period also exhibited seasonal patterns, with most major storms occurring in the hot season, with a peak between February and May (Figure \ref{figure_cyclone}).

@@

```{r figure_gps_climate, include = TRUE, echo = FALSE, warnings = FALSE, fig.scap="Spatial distribution of climate conditions across clinics and seasons in Madagascar.", out.extra='', fig.cap="\\label{figure_gps_climate}Spatial distribution of climate conditions across clinics and seasons in Madagascar. Mean monthly temperature (top) is measured in degrees Celsius  and mean soil moisture is measured in percent saturation (below).", fig.width = 6, fig.height = 7.5, fig.align='center', out.width = '90%'}

@@

The regression models predicting monthly counts of any mental or neurological disorder (Table 2.2) showed significant associations with temperature across all models. A one-degree Celsius increase in temperature was associated with approximately `r function_dynamic_text_results(section = "main", model_option = 1, coefficient_option = "exposure_temperature", value = "irr", value_include ="only")` visits reported `r function_dynamic_text_results(section = "main", model_option = 1, coefficient_option = "exposure_temperature", value = "irr", value_include =TRUE)` by primary clinics and hospitals in Model 1. Inclusion of both climate predictors resulted in little change in the measure of association for temperature. In an additive model with soil moisture included (Model 3), the effect estimate associated with temperature grew to `r function_dynamic_text_results(section = "main", model_option = 3, coefficient_option = "exposure_temperature", value = "irr", value_include ="only")` visits reported `r function_dynamic_text_results(section = "main", model_option = 3, coefficient_option = "exposure_temperature", value = "irr", value_include =TRUE)`. Adopting an interactive model (Model 4) attenuated the association `r function_dynamic_text_results(section = "main", model_option = 4, coefficient_option = "exposure_temperature", value = "irr", value_include =TRUE)`, but remained significant at a 95% level of confidence. Across all models with temperature, the average marginal effect associated with temperature (across a sample with all clinics) ranged between 2 and 3 fewer monthly mental health visits reported and was statistically significant at an alpha of 0.01, with the average marginal effect in Model 4 being `r function_dynamic_text_results(section = "main", model_option = 4, coefficient_option = "exposure_temperature", value = "ame", value_include ="only")` visits `r function_dynamic_text_results(section = "main", model_option = 4, coefficient_option = "exposure_temperature", value = "ame", value_include =FALSE)`.

@@

```
Predicted values of monthly mental health related visit counts under Model 4 echoed the effect estimates from Table 2, with fewer visits reported in warmer months (Figure \ref{figure_marginal}). Marginal effect estimates across the distribution of temperatures in our data showed that the greatest level of reduction occurred between 10 and 15 degrees Celsius, with each degree increase associated with 4 to 5 fewer visits related to mental or neurological disorders reported, controlling for other covariates. Months with greater levels of soil moisture and days with cyclone activity also showed fewer reported visits; however, the marginal effects associated with soil moisture were not significant across any unit of exposure, while those associated with days with cyclone activity were close to zero for all lengths of activity.

Examining lagged effects showed potential temporal associations between soil moisture levels in preceding months and visit counts in those following (Figure \ref{figure_lag}), with higher saturation leading to a statistically significant elevation in visits reported 1 to 5 months later. At a lag of 3 months, the elevation in visits was approximately `r function_dynamic_text_results(section = "lags", lag_option = 3, coefficient_option = "exposure_sm", value = "irr", value_include = "only")` `r function_dynamic_text_results(section = "lags", lag_option = 3, coefficient_option = "exposure_sm", value = "irr", value_include =TRUE)`. There appeared to be a significant elevation of similar magnitude 13 to 17 months prior, suggesting a potential annual pattern in the lagged association. Effects for temperature were not significant across any lag; cyclone activity similarly showed no significance in lags under a year, but did show modest, statistically significant associations at lags of 12 months `r function_dynamic_text_results(section = "lags", lag_option = 12, coefficient_option = "exposure_cyclone", value = "irr", value_include = TRUE)` and 15 months `r function_dynamic_text_results(section = "lags", lag_option = 15, coefficient_option = "exposure_cyclone", value = "irr", value_include = TRUE)`. A slight increase in visits was associated with cyclone activity in a prior month, but this increase was not statistically significant `r function_dynamic_text_results(section = "lags", lag_option = 1, coefficient_option = "exposure_cyclone", value = "irr", value_include = TRUE)`.

@@

```{r figure_sensitivity_analysis_temp, include = TRUE, echo=FALSE, warnings = FALSE, fig.scap=paste("Specification curve depicting the effect estimates and confidence intervals for mean monthly temperature (degrees Celsius) across", num_models_multiverse, "models."), out.extra='', fig.cap=paste("\\label{figure_sensitivity_analysis_temp}Specification curve depicting the effect estimates and confidence intervals for mean monthly temperature (degrees Celsius) across", num_models_multiverse, "models. In the top panel, estimated incidence rate ratios (IRRs) across different models are shown as circles; the darkly and lightly shaded regions above and below these circles correspond to 95\\% and 99\\% confidence intervals of these estimates, respectively. The dotted horizontal line depicts a null effect. The IRR estimate and confidence intervals of the preferred model, Model 4, are highlighted in orange. Below each model, a legend of shaded and unshaded boxes indicates which modeling decisions correspond to a given model, with boxes shaded teal showing active features. FE: fixed effects. CSB/FSP: Centres Santé de Base and Formations Sanitaires Privées de Base. NB: negative binomial model.", sep=" "), fig.width = 15, fig.height = 9, fig.align='center', out.width = '95%'}

@@

```{r figure_sensitivity_analysis_sm, include = TRUE, echo=FALSE, warnings = FALSE, fig.scap=paste("Specification curve depicting the effect estimates and confidence intervals for mean monthly soil moisture (percent saturation) across", num_models_multiverse, "models."), out.extra='', fig.cap=paste("\\label{figure_sensitivity_analysis_sm}Specification curve depicting the effect estimates and confidence intervals for mean monthly soil moisture (percent saturation) across", num_models_multiverse, "models. In the top panel, estimated incidence rate ratios (IRRs) across different models are shown as circles; the darkly and lightly shaded regions above and below these circles correspond to 95\\% and 99\\% confidence intervals of these estimates, respectively. The dotted horizontal line depicts a null effect. The IRR estimate and confidence intervals of the preferred model, Model 4, are highlighted in orange. Below each model, a legend of shaded and unshaded boxes indicates which modeling decisions correspond to a given model, with boxes shaded teal showing active features. FE: fixed effects. CSB/FSP: Centres Santé de Base and Formations Sanitaires Privées de Base. NB: negative binomial model.", sep=" "), fig.width = 15, fig.height = 9, fig.align='center', out.width = '95%'}

@@

```{r figure_sensitivity_analysis_cyclone, include = TRUE, echo=FALSE, warnings = FALSE, fig.width = 15, fig.height = 9, fig.align='center', out.width = '95%', fig.scap=paste("Specification curve depicting the effect estimates and confidence intervals for duration of cyclone activity across", num_models_multiverse, "models."), out.extra='', fig.cap=paste("\\label{figure_sensitivity_analysis_cyclone}Specification curve depicting the effect estimates and confidence intervals for cyclone activity (measured as days with a tropical storm or cyclone impacting Madagascar in a month) across", as.numeric(num_models_multiverse)/3, "models. In the top panel, estimated incidence rate ratios (IRRs) across different models are shown as circles; the darkly and lightly shaded regions above and below these circles correspond to 95\\% and 99\\% confidence intervals of these estimates, respectively. The dotted horizontal line depicts a null effect. The IRR estimate and confidence intervals of the preferred model, Model 4, are highlighted in orange. Below each model, a legend of shaded and unshaded boxes indicates which modeling decisions correspond to a given model, with boxes shaded teal showing active features. FE: fixed effects. CSB/FSP: Centres Santé de Base and Formations Sanitaires Privées de Base. NB: negative binomial model.", sep=" ")}

@@

It is important to note that the true relationship between reported mental health visits and actual cases of mental disorders is not well understood in Madagascar due to the lack of epidemiological data on the prevalence or distribution of mental disorders in the general population. The changes in mental health visits could reflect changes in the actual cases of mental disorders but could also result from the influence of \acr{CCEs} on help-seeking behavior, diagnostic capacity, or other factors. Additionally, it has been widely recognized that the majority individuals with mental disorders often do not receive any mental health treatment, particularly in low-income settings. [@demyttenaerePrevalenceSeverityUnmet2004]

@@

An alternative explanation could be that access to mental health services is limited in warmer months due to factors such as the increased demand for medical services related to heat-related illnesses. A crowding out of provider availability due to physical complaints in warmer months (for example, malaria-related fever) could lead to decreased availability of provider access for mental disorders, especially at primary health facilities with already limited diagnostic capacity. Understanding how climate conditions in Madagascar impact the broader burden of disease, including physical conditions, is the subject of current investigation, and will lend further insight into potential mechanistic explanations for our findings.

@@

Though incidence rate ratios and marginal effect estimates associated with soil moisture were not statistically significant in our main analysis, we observed a significant association of higher soil moisture and visits for mental health complaints in the following quarter, echoing similar findings from United States meteorological data linking higher precipitation with worsened mental health.[@obradovichEmpiricalEvidenceMental2018] In addition, our subgroup analysis also indicated significant and divergent effects among different ecoregions. In particular, we found that greater soil moisture was associated with more visits reported in the eastern lowlands and the northern dry deciduous forests, which echoes the geographic concentration of flood risk in Madagascar.[@rakotoarisonSpatialAnalysisRisks2021] Conversely, the observed (though statistically modest) reduction in visits associated with higher soil moisture in the desert and xeric shrublands overlaps with the concentration of drought risk, where higher soil moisture could indicate reduced drought vulnerability. It is important to note, however, that the time frame of available GESIS data excludes the most severe period of the ongoing drought in Madagascar, limiting our ability to directly study the impact of the current crisis on population mental health in this analysis. Further research could leverage spatial reanalysis data on precipitation to further characterize these associations.
